---
title: "Homework#2"
author: "Elise Rust"
date: "9/22/2021"
output: html_document
---

<style type="text/css">
  body{
  font-size: 12pt;
  font-family: "Times New Roman", Times, serif;
}
</style>

Set seed and load in necessary packages
```{r}
set.seed(225)

library(tidyverse)
library(ggplot2)
library(stats)
library(nortest)
```

# Problem #1:
Demonstrate that these two commands indeed give the same result for 5 diff values of p (use table function)

Initializing myattempts(p) function and rgeom(1,p).
```{r}
## Random variable defined by counting the number of failures until the first succcess

# myattempts(p)
mytoss = function(p){
  u <- runif(1)
  x <- as.numeric(u < p)
  return(x)
}

myattempts = function(p) { counter <- 0
while (mytoss(p) == 0) { counter <- counter + 1 }
return(counter) }
```
### I answered parts b - e.

b. For the second p, compare distribution by computing statistics such as mean and standard deviation
```{r}
p = 0.28

method1 = replicate(100000, myattempts(p))
method2 = replicate(100000, rgeom(1,p))

# compute statistics for each
summary_stats1 <- summary(method1)
summary_stats2 <- summary(method2)

# Compute standard deviation for each
standard_dev1 <- sd(method1)
standard_dev2 <- sd(method2)

# Print
cat("The mean of the random variable defined by myattempt(0.28) is: ", summary_stats1[4], ", its median is: ", summary_stats1[3], ", its maximum is: ", summary_stats1[6], ", and its standard deviation is: ", standard_dev1, ".")
cat("The mean of the random variable defined by rgeom(1,0.28) is: ", summary_stats2[4], ", its median is: ", summary_stats2[3], ", its maximum is: ", summary_stats2[6], ", and its standard deviation is: ", standard_dev2, ".")
```

c. For the third p, plot both the distributions as histograms in the same plot
```{r}
p = 0.55

method1 = replicate(100000, myattempts(p))
method2 = replicate(100000, rgeom(1,p))

# convert into dataframes in order to plot
df1 <- as.data.frame(method1)
df2 <- as.data.frame(method2)

# plot both the distributions as histograms on the same plot
ggplot(df1, aes(x=method1)) + 
  geom_histogram(bindwidth = 1, alpha=0.5, fill = "pink") +
  geom_histogram(data = df2, aes(x = method2, bindwidth = 1, alpha = 0.5, fill = "white")) +
  labs(title = "Comparative histograms of distributions of myattempts() and rgeom()")
```

d. For the fourth p, make side-by-side box plots
```{r}
p = 0.91

method1 = replicate(100000, myattempts(p))
method2 = replicate(100000, rgeom(1,p))

# convert into dataframes in order to plot
df1 <- as.data.frame(method1)
df2 <- as.data.frame(method2)

head(df1)
head(df2)

# combine dataframes and gather them
df_big = cbind(df1, df2)
df_big <- gather(df_big, "type", "value", 1:2)

head(df_big)

# plot side-by-side box plots
ggplot(df_big, aes(x= type, y= value, fill=type)) +
  geom_boxplot() +
  theme_classic()
```

e. For the fifth p, plot the two empirical distribution functions in the same plot.
```{r}
p = 0.63

method1 = replicate(100000, myattempts(p))
method2 = replicate(100000, rgeom(1,p))

# plot two empirical distribution functions in the same plot
require(graphics)
plot(ecdf(method1), verticals = TRUE, do.points = FALSE, col = 'green')
plot(ecdf(method2), verticals = TRUE, do.points = FALSE, add = TRUE, col = 'orange')
```

# Problem #2:

```{r}
# Draw a uniformly distributed random number X1 from the interval (0,1)
number_generator <- function() {
  X1 = runif(1, min = 0, max = 1)
  X2 = runif(1, min = 0, max = 1 + X1)
  X3 = runif(1, min = 0, max = 1 + X2)
  X4 = runif(1, min = 0, max = 1 + X3)
  X5 = runif(1, min = 0, max = 1 + X4)
  X6 = runif(1, min = 0, max = 1 + X5)
  X7 = runif(1, min = 0, max = 1 + X6)
  X8 = runif(1, min = 0, max = 1 + X7)
  X9 = runif(1, min = 0, max = 1 + X8)
  X10 = runif(1, min = 0, max = 1 + X9)
  
  return(X10)
}

# Use a monte carlo simulation to give an approximate answer to: What is the mean value of X10?
# Run 10,000 times
runs <- 10000
simulations <- replicate(runs, number_generator())

mean_X10 <- print(mean(simulations))
cat("Using a Monte Carlo Simulation that runs 10,000 times we find that the mean value of X10 is: ", mean_X10)

# Use a histogram to identify the distribution of X10
hist(simulations, freq = TRUE, col = "pink", main = "Histogram of the distribution of X10 after 10,000 simulations.", 
     xlab = "Value of X10", ylab = "Frequency")

```

# Problem #3: X has a gamma distribution with shape parameter r = 2.5 and scale parameter p = 5.

a. Prob(X <= 10)
```{r}
shape = 2.5
scale = 5
x = 10

# Use gamma distribution --> cumulative distribution for X <= 10
prob1 = pgamma(x, shape, scale, lower.tail = TRUE)
cat("The probability that random variable X - which is distributed via the Gamma distribution - is less than or equal to 10 equals: " , prob1)
```
b. Prob(X > 5)
```{r}
# Prob(X > 5) = 1 - Prob(X <= 4)
shape = 2.5
scale = 5
x = 4

# Use gamma distribution --> cumulative distribution for 1 - X <= 4
prob1 = pgamma(x, shape, scale, lower.tail = FALSE)
cat("The probability that random variable X - which is distributed via the Gamma distribution - is greater than 5 equals: " , prob1)
```

c. Prob(|X - 8| < 1)
(|X - 8|) = 7 < X < 9
```{r}
shape = 2.5
scale = 5
x1 = 9
x2 = 8
# P(X < 9) & P(X > 7) = P(X < 9) - (1 - P(X < 8))

# Use gamma distribution
prob1 = pgamma(x1, shape, scale) - (1 - pgamma(x2, shape, scale))
cat("The probability that random variable X - which is distributed via the Gamma distribution - is between 7 and 9 equals: " , prob1)
```

d. z such that Prob(X < z) = 0.95
```{r}
# Use quantile distribution
p <- 0.95

z = qgamma(p, shape, scale)

cat("Value z such that P(X < z) = 0.95 is: ", z)

# Double check that it worked
#pgamma(1.10705, shape, scale)
```


# Problem #4:
Is binomial distribution B(n,p) close to that of a normal distribution?

Plot cumulative distribution functions in the same figure using a staircase plot for binomial and a line plot for normal
a. Case 1: np and n(1-p) are large
```{r}
x = 0:200
n = 200
p = 0.9
mean = mean(x)
sd = sd(x)

binomial <- pbinom(x, size = n, prob = p)
normal <- pnorm(x, mean, sd)

plot(x, binomial, type = "S")
plot(x, normal, type = "l")
```

The binomial distribution is steeper and skewed more right than the normal distribution. The normal distribution's cdf is a smooth upward sloping line between 0 and 0.8 while the binomial distribution's cdf is a constant line until around 175 where it sky rockets up to 1. However, relative to the other two cases it resembles a normal distribution far more.

b. np is large and n(1-p) < 10
```{r}
x = 0:15
n = 15
p = 0.8 
# np = 12, n(1-p) = 3
mean = mean(x)
sd = sd(x)

binomial <- pbinom(x, size = n, prob = p)
normal <- pnorm(x, mean, sd)

plot(x, binomial, type = "S")
plot(x, normal, type = "l")
```

In this case the binomial staircase plot is less smooth and resembles a normal distribution less. The smaller n means the plot is less normalized and increments in larger chunks.

c. np < 10 and n(1 − p) < 10
```{r}
x = 0:10
n = 10
p = 0.8 
# np = 8, n(1-p) = 2
mean = mean(x)
sd = sd(x)

binomial <- pbinom(x, size = n, prob = p)
normal <- pnorm(x, mean, sd)

plot(x, binomial, type = "S")
plot(x, normal, type = "l")
```

This plot is similar to plot 2, but the effect is worsened as n is again even smaller. Here, neither np nor n(1-p) are greater than 10 so it resembles a normal distribution the least.

# Problem #5:
```{r}
# sample size of 10
sample10 <- rnorm(10)
qqnorm(sample10)
```

Sample size 10: The plot is the farthest from a straight line of the group. Though it slightly resembles a straight line, with this few samples it is hard to tell accurately whether the sample comes from a normal distribution.

```{r}
# sample size of 20
sample20 <- rnorm(20)
qqnorm(sample20)
```

Sample size 20: With more samples (i.e. more points on the graph) this plot begins to resemble a straight line more. We can begin to see that this sample is likely normally distributed.

```{r}
# sample size of 40
sample40 <- rnorm(40)
qqnorm(sample40)
```

Sample size 40: This sample is even more normally distributed as it more closely resembles a straight line than the previous two. Between the -1 and 1 theoretical quantiles is where the line is the straighest whereas as the plot gets farther from 0 it is more scattered.

```{r}
# sample size of 100
sample100 <- rnorm(100)
qqnorm(sample100)
```

Sample size 100: This sample continues the trend and again resembles a straight line more than the previous three plots did. With this number of values it is pretty clear that the sample is normally distributed and most closely resembles a straight line between the -1 and 1 theoretical quantiles.

```{r}
# sample size of 400
sample400 <- rnorm(400)
qqnorm(sample400)
```

Sample size 400: Finally, with a sample size of 400 values the sample clearly has an approximate normal distribution. This time, more of the plot resembles a straight line (i.e. from the -2 to 2 theoretical quantiles) and the scattering at the tails is tighter.


#Problem #6:

a. Simulation #1 --> Standard Normal Distribution
```{r}
### X has a continuous distribution
### f(x) = pdf
### F(X) = cdf --> cumulative distribution function
### U = F(X) has a uniform U(0,1) distribution

# Generate random sample - normally distributed
X = rnorm(100)

# plugging it into cdf
CDF_F= ecdf(X)

# sort data
U = CDF_F(sort(X))

# plot U
plot(U, main = "Distribution of U = F(X) of normally distributed random variable X")

# Check with Kolmogorov-Smirnov Test
ks.test(U, "punif")

#################
# Alternative code I was working with that may also be correct:
# X = rnorm(100)
#U = pnorm(X, mean = 0, sd = 1)
#y = seq(-10, 9.8, by = 0.2)
#plot(y, U,  type = "l", main = "CDF of the Normal Distribution")
```

b. Simulation #2 --> Exponential Distribution
```{r}
# Generate ranom varible - Exponentially distributed
X = rexp(100)

# plugging it into cdf F
CDF_F = ecdf(X)

# sort data
U = CDF_F(sort(X))

# plot U
plot(U, main = "Distribution of U = F(X) of exponentially distributed random variable X")


## Again, the alternative I was playing with that uses pexp instead
#X = rexp(100)
# U = pexp(X)
# y = seq(0,99, by = 1)
#plot(X, U,  main = "CDF of the Exponential Distribution")
```

c. Simulation #3 --> Gamma Distribution
```{r}
# Generate ranom varible - Gamma distributed
X = rgamma(100, shape = 2.5, scale = 5)

# plugging it into cdf F
CDF_F = ecdf(X)

# sort data
U = CDF_F(sort(X))

# plot U
plot(U, main = "Distribution of U = F(X) of Gamma distributed random variable X")


## Again, the alternative solution I was playing with
#X = rgamma(100, shape = 2.5, scale = 5)
# U = pgamma(X, shape = 2.5, scale = 5)
# y = seq(from = 0, to = 99, by = 1)
# plot(x, U,  main = "CDF of the Gamma Distribution")
```

# Bonus Problem:
```{r}
# Generate X1 and X2
N = 10000
X1 = rexp(N)
X2 = rexp(N)

# Sum X1 and X2
X = X1 + X2

# Use qqnorm plots to determine an appropriate value of alpha
plot(qqnorm(X))

# Now use a loop to test for alphas that would make a normally distributed X^a
suitable = FALSE
i = 1

while (suitable == FALSE) {
  test_X = X^i
  # Run Adnerson test
  result = ad.test(rnorm(test_X))
  
  if (result [2] > 0.05) {
    suitable = TRUE
    print("An approximate value for alpha is: ")
    print(i)
  }
  else {
    i = i + 1
  }
}

```