---
title: "HW3_Rust"
author: "Elise Rust"
date: "10/12/2021"
output: html_document
---

<style type="text/css">
  body{
  font-size: 12pt;
  font-family: "Times New Roman", Times, serif;
}
</style>

Load in Libraries
```{r}
library(dplyr)
library(tidyverse)
```

# Problem #1: True or False - Random Walk

False. The transition matrix is not symmetric because some tiles have more subsequent tiles to move to whereas others (along the corners and edges) have fewer tile choices for the next move. For example, if moving from tile 1 on X1 to tile 2 on X2, $P(X_{i+1} = 2| X_i = 1)$ = 0.5 because one can only move to to two tiles from tile 1. However, if moving from tile 2 to tile 1, $P(X_{i+1} = 1|X_i = 2)$ = 0.33. Therefore, these probabilities do not equal each other and the transition matrix is not symmetric.


# Problem #2
```{r}
size = 10000

X = rbinom(size, 80, 0.2)
Y = rbinom(size, 100, 0.7)

Z = X + Y
```

a) P(X < 12|X < 18) and E(X|X < 18)
```{r}
sub_18 <- X[X < 18]

# P(X < 12 | X < 18)
sub_12 <- sub_18[sub_18 < 12]
prob1 = length(sub_12) / length(sub_18)

cat("The conditional probability of X (from binomial distribution B(80, .2)) less than 12 given X < 18 is: ", prob1)

# E(X } X < 18)
cat("The expected value of X|X<18 is: ", mean(sub_18))
```

b) Cumulative distribution function of X|(12 <= X <= 20)
```{r}
subset2 <- between(X, 12, 20)
Xecdf <- ecdf(subset2)
t <- seq(0, 2, by = 0.1)
plot(t, Xecdf(t), type = "l", lwd = 2)
```

c) Cumulative distribution function of X|Z = 90
```{r}
Z90 = Z[Z == 90]

Xecdf <- ecdf(Z90)
t <- seq(0, 2, by = 0.1)
plot(t, Xecdf(t), type = "l", lwd = 2)
```


d) E(Z|X=k)fork=10,15,20
```{r}
EV10 = mean(Z[X == 10])
EV15 = mean(Z[X == 15])
EV20 = mean(Z[X == 20])

cat("The expectation of variable Z given an X value of 10 is: ", EV10)
cat("The expectation of variable Z given an X value of 15 is: ", EV15)
cat("The expectation of variable Z given an X value of 20 is: ", EV20)
```

e) E(X|Z = k) for k = 80,90,100
```{r}
EV80 = mean(X[Z == 80])
EV90 = mean(X[Z == 90])
EV100 = mean(X[Z == 100])

cat("The expectation of variable Z given an X value of 10 is: ", EV10)
cat("The expectation of variable Z given an X value of 15 is: ", EV15)
cat("The expectation of variable Z given an X value of 20 is: ", EV20)
```



# Problem #3: 
```{r}
# Initialize X and Y
lambda = 1
N = 10000

df3 = as.data.frame(rexp(10000, lambda)) # Calculate X
names(df3) <- c("X") # rename column

df3 <- df3 %>%
  mutate(Y = rpois(10000, X)) # Calculate Y
```

a) Generate at least 1000 random samples from the marginal distribution of X and make a probability histogram.
```{r}
samples = 1000

hist(replicate(1000, df3$X), prob = T)
```

b) Generate at least 1000 random samples from the conditional distribution of Y |X = 1.5 and make a probability histogram.
```{r}
#df3$Y[df$X == 1.5] # No results, X never equals 1.5

## Use round function
df3$rounded_x <- round(df3$X, digits = 2) # round X to 2 digits
head(df3) # view results

hist(replicate(1000, df3$Y[df3$rounded_x == 1.5]), prob = T)
```

c) Generate at least 1000 random samples from the marginal distribution of Y and make a probability histogram.
```{r}
hist(replicate(1000, df3$Y), prob = T)
```

d) Generate at least 1000 random samples from the conditional distribution of X|Y = 2 and make a probability histogram.
```{r}

hist(replicate(1000, df3$X[df3$Y == 2]), prob = T)
```


# Problem #4: 
Suppose X and Y have independent standard normal distributions. Make at least 1,000 random samples from Z, defined as Z = Y |(X + Y ≥ 1). Do you think that Z has a normal distribution? What are its approximate mean and standard deviation?
```{r}
X = rnorm(1000, 0, 1)
Y = rnorm(1000, 0, 1)

Z = Y[X + Y >= 1]

plot(ecdf(Z))

mean(Z)
sd(Z)
```
Based on the above results, the ecdf of Z does resemble the known ecdf of a normal distribution while the mean is around 1. The standard deviation is 0.7739997 though, so this is not a standard normal distribution. It is however largely resembling a normal distribution.

# Problem #5: 

1) Written response on attached PDF

4) Generate a sample of size 10,000 from Y1 ∼ N(−2,1), Y2 ∼ N(2,2), w1 = 1/3, w2 = 2/3 and make a
probability histogram. Clearly this is not a normal distribution, and a mixture is not a sum!
```{r}
N = 10000
Y1 = rnorm(N, -2, 1)
Y2 = rnorm(N, 2, 2)

w1 = 1/3
w2 = 2/3

X = as.array(sample (c(1,2), size=10000, replace =T)) # define X

# Create a dataframe
Y1 = as.array(Y1)
Y2 = as.array(Y2)
df4 = as.data.frame(cbind(Y1, Y2, X))

# Create variable Y
df4 <- df4 %>%
  mutate(Y = case_when(
    X == 1 ~ Y1,
    X == 2 ~ Y2
  ))

# Plot the mixture of Y
hist(df3$Y, prob = T)
```
Definitely not a normal distribution!