---
title: "HW4_Rust"
author: "Elise Rust"
date: "10/28/2021"
output: html_document
---

<style type="text/css">
  body{
  font-size: 12pt;
  font-family: "Times New Roman", Times, serif;
}
</style>

```{r}
library(tidyverse)
library(ggplot2)
library(plotrix)
```

# Problem #1: 
```{r}
X = rmultinom(n = 10, size = 10, prob = c(.2, .5, .3))
print(X)
X = as.data.frame(X)

count = 0

for (i in 1:ncol(X)) {
  result <- (X[,i] == c(2,4,4))
  result = as.vector(result)
  #print(result)

  if (result[1] == "TRUE") {
    if (result[2] == "TRUE") {
      if (result[3] == "TRUE") {
        count = count + 1
        print(count)
      }
    }
  }
}

#print(count)

prob = count/length(X)
cat("The probability that multinomial X has X1 = 2, X2 = 4, and X3 = 4 is: ", prob)

# Check using dmultinom()
x = c(2, 4, 4)
dmulti = dmultinom(x, size = 10, prob = c(.2, .5, .3))
cat("Using dmultinom(), we find that the probability that X1 = 2, X2 = 4, and X3 = 4 is equal to: ", dmulti)
```

# Problem #2:
```{r}
# Use sample() first
# Generate X dataframe
X_df <- function() {
  X1 = sample(0:1, size = 10, prob = c(0.8, 0.2), replace = TRUE)
  table = as.data.frame(table(X1))
  X1_count = table$Freq[table$X1 == '1']
  #print(X1_count)
  
  X2 = sample(0:1, size = 10, prob = c(0.5, 0.5), replace = TRUE)
  table = as.data.frame(table(X2))
  X2_count = table$Freq[table$X2 == '1']
  #print(X2_count)
  
  X3 = sample(0:1, size = 10, prob = c(0.7, 0.3), replace = TRUE)
  table = as.data.frame(table(X3))
  X3_count = table$Freq[table$X3 == '1']
  #print(X3_count)
  
  X = c(X1_count, X2_count, X3_count)
  return(X)
}

X = as.data.frame(replicate(10, X_df()))

count = 0

for (i in 1:ncol(X)) {
  result <- (X[,i] == c(2,4,4))
  result = as.vector(result)
  #print(result)

  if (result[1] == "TRUE") {
    if (result[2] == "TRUE") {
      if (result[3] == "TRUE") {
        count = count + 1
        print(count)
      }
    }
  }
}

print(count)

prob = count/length(X)
cat("The probability that multinomial X has X1 = 2, X2 = 4, and X3 = 4 is: ", prob)

# Check using dmultinom()
x = c(2, 4, 4)
dmulti = dmultinom(x, size = 10, prob = c(.2, .5, .3))
cat("Using dmultinom(), we find that the probability that X1 = 2, X2 = 4, and X3 = 4 is equal to: ", dmulti)
```


# Problem #3:
```{r}
sim <- function() {
  X = runif(12, 0, 1)
  X = as.data.frame(X)
  colnames(X) = c("X_values")
  
  Z = sum(X$X_values)  - 6 # set up Z
  return(Z)
  #print(Z)
}

Z10 <- replicate(10, sim())
Z100 <- replicate(100, sim())
Z1000 <- replicate(1000, sim())
Z10000 <- replicate(10000, sim())

plot10 <- plot(ecdf(Z10))
plot100 <- plot(ecdf(Z100))
plot1000 <- plot(ecdf(Z1000))
plot10000 <- plot(ecdf(Z10000))

cat("The mean of Z for n large isL ", mean(Z10000), " and the standard deviation for n large is: ", sd(Z10000), " ; therefore, Z = X1 + X2 + ... + X12 - 6 has an approximate standard normal distribution.")
```


# Problem #4:
a) Give the exact sampling distribution
(see attached PDF for theoretical sampling distribution)

b) Simulate the sampling distribution of W in R and plot the results.
```{r}
X_fxn <- function() {
  X_i = rnorm(100, mean = 7, sd = 9)
}
Y_fxn <- function() {
  Y_i = rnorm(100, mean = 10, sd = 5)
}

W_fxn <- function() {
  X = as.data.frame(replicate(9, X_fxn()))
  Y = as.data.frame(replicate(12, Y_fxn()))
  
  X_mean = mean(as.matrix(X))
  Y_mean = mean(as.matrix(Y))
  
  W = X_mean - Y_mean
  return(W)
}

W_1000 = replicate(10000, W_fxn())
cat("The mean of the simulated distribution W = mean(X) - mean(Y) for 10,000 samples is: ", mean(W_1000))
cat("The variance of the simulated distribution W = mean(X) - mean(Y) for 10,000 samples is: ", sd(W_1000)*sd(W_1000))

plot(ecdf(W_1000))
```

c) Find probability of P(W<-1.5)
```{r}
P_1.5 = W_1000[W_1000 < -1.5]
prob = length(P_1.5)/length(W_1000)
cat("The probability that P(W < -1.5) is: ", prob)
```

# Problem #5
a) Simulate the sampling distribution of X-mean
```{r}
sim <- function() {
  X_i = rexp(30, rate = 1/3)
  X_mean = mean(X_i)
  return(X_mean)
}

cat("The sampling distribution of X-mean is: ", sim())

X_df = as.data.frame(replicate(1000, sim()))
names(X_df) <- c('values')

X_df %>% 
  ggplot(aes(x = values)) +
  geom_density()
```

b) Find the mean and standard error of the sampling distribution and compare to the theoretical results
```{r}
cat("The mean of the sampling distribution is: ", mean(X_df$values), " , and the standard error is: ", std.error(X_df$values))
```

c) From your simulation, find P(X <= 3.5)
```{r}
X_3.5 = as.data.frame(X_df[X_df <= 3.5])
prob = nrow(X_3.5)/nrow(X_df)
cat("The probability that X mean is less or equal to 3.5 is: ", prob)
```


d) Estimate P(X <= 3.5) by assuming that the CLT approximation holds. Compare to part (C).
```{r}
Z = rnorm(10000, mean(X_df$values), sd(X_df$values))
Z = as.data.frame(Z)
Z_3.5 = as.data.frame(Z[Z <= 3.5])
prob = nrow(Z_3.5)/nrow(Z)
cat("The probability that X mean is less or equal to 3.5 is: ", prob)
```

The probabilities are essentially the same! Using the original simulation we get a probability of 0.816, and using the Central Limit Theorem we get a probability of 0.817 - indicating that the assumption of CLT approximation is valid.